ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-02-26 pyked.tex
edited discussion of PrIMe flaws a bit

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} represents an admirable \todo{Maybe its just me, but this feels a bit sarcastic} first attempt at standardization, but a number of limitations
prevent wide adoption: the PrIMe standard does not require or support all information
needed to simulate an experiment, including a machine-readable definition of ignition or a
standard way to express the rapid compression machine (RCM) compression process; and
the XML format, intended more as a markup language than data format, presents a barrier
to creating and working with files.
Furthermore, the closed and opaque nature of the PrIMe standard and associated Data Warehouse
limit contributions and community development of these and compatible tools.

More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the ReSpecTh standard,
an XML-based format that builds on PrIMe. ReSpecTh adds important features that make files
better standalone representations of experimental data---i.e., more informative by themselves
rather than in concert with a larger system.
For example, where PrIMe used internal bibliographic references, ReSpecTh added a field for
typical bibliographic data. The ReSpecTh standard also provides machine-readable formats
for better specifying ignition experiments, including a definition of autoignition and
either volume histories for RCMs or pressure-rise in shock tubes.
However, being XML-based, ReSpecTh experimental files still suffer from the same usability
issues as PrIMe, and the standard remains closed.
\todo[author=Kyle,inline]{how's this?}

Olm and colleagues used their ReSpecTh-based database to quantify the performance of literature
hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models, and more recently
develop improved models for hydrogen\slash syngas~\autocite{Varga:2016gj} and
ethanol~\autocite{Olm:2016et}. In all four cases, they converted numerous experimental datasets
from the literature into the ReSpecTh standard.

>>>>

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make it
a strong standardization format, but suffers from two primary flaws that
prevent wide adoption: (1) the PrIMe standard does not require or support all information
needed to simulate an experiment, including a machine-readable definition of ignition or a
standard way to express the rapid compression machine (RCM) compression process; (2) and
the XML format, intended more as a markup language than data format, presents a barrier
to creating and working with files.
Furthermore, the focus on using internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for scholarly products
and InCHI or SMILES for species), limits usability outside the PrIMe ecosystem.
Finally, the closed and opaque nature of the PrIMe standard and associated Data Warehouse
limit contributions and community development of these and compatible tools.

More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the ReSpecTh standard,
an XML-based format that builds on PrIMe. ReSpecTh adds important features that make files
better standalone representations of experimental data---i.e., more informative by themselves
rather than in concert with a larger system.
For example, where PrIMe used internal bibliographic references, ReSpecTh added a field for
typical bibliographic data. The ReSpecTh standard also provides machine-readable formats
for better specifying ignition experiments, including a definition of autoignition and
either volume histories for RCMs or pressure-rise in shock tubes.
However, being XML-based, ReSpecTh experimental files still suffer from the same usability
issues as PrIMe, and the standard remains closed.
Olm and colleagues used their ReSpecTh-based database to quantify the performance of literature
hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models, and more recently
develop improved models for hydrogen\slash syngas~\autocite{Varga:2016gj} and
ethanol~\autocite{Olm:2016et}. In all four cases, they converted numerous experimental datasets
from the literature into the ReSpecTh standard.
\todo[author=Kyle,inline]{how's this?}
